---
title: "Local weather"
author: "Yves"
date: "18 August 2024"
output:   
  flexdashboard::flex_dashboard:
    orientation: columns
runtime: shiny
---

```{r setup, include=FALSE}

library(plotly)
library("RColorBrewer")
library(lubridate)
library(kableExtra)
library(httr2)
library(readr)
library(stringr)
library(tidyverse)
library(sf)
library(leaflet)

RV <- reactiveValues()

source("C:\\Users\\ylace\\OneDrive\\Desktop\\STUFF\\Weather\\GET_historical_weather.R")
source("C:\\Users\\ylace\\OneDrive\\Desktop\\STUFF\\Weather\\GET_forecast_weather.R")
#source("C:\\Users\\ylace\\OneDrive\\Desktop\\STUFF\\Weather\\GET_world_weather.R")

init <- function() {
  lat <- 39.48083560687681
  lon <- -105.06391035422446
  address <- "8204 Mount Kataka St, Littleton, CO 80125"
  date <- Sys.Date()
  df <- data.frame(address, lat, lon, date)
  return(df)
}
location <- init()
RV$init <- location

point <- c(location$lat, location$lon)
RV$url <- GET_forecast_url(point)

```

<b></b> {data-icon="ion-stats-bars"}
=====================================  

Column {data-width=100}
-------------------------------------
### <b></b>

```{r, echo=FALSE}
  actionButton("reset1", "Map")
```

```{r, echo=FALSE}
observeEvent(input$reset1,{
    proxy <- leafletProxy("EC")
    proxy %>% addTiles() %>% 
     addProviderTiles("CartoDB.Positron") %>% 
     setView(lng=location$lon, lat=location$lat, zoom = 10) %>% 
     addPolygons(data=DC.CO, color = "green", 
                 label="Douglas County, CO", fillOpacity=.1) %>%
     addCircles(lng=location$lon, lat=location$lat, 
                radius=520, color='red', 
                label=location$address) %>%
     addCircles(lng=KASSLER$longitude, lat=KASSLER$latitude, 
                color='black', radius=520,
                label=paste(KASSLER$station, KASSLER$name)) %>%
     addCircles(lng=stations$longitude, lat=stations$latitude, 
                label=paste(stations$station, stations$name))
})
```

Column {}
-------------------------------------

### <b>Weather stations in and around Douglas County, CO</b>
```{r, echo=FALSE}
output$EC <- renderLeaflet({
   leaflet() %>% 
     addProviderTiles("CartoDB.Positron") %>% 
     addPolygons(data=DC.CO, color = "green", 
                 label=DC.CO$STATEFP, fillOpacity=.1) %>%
     addCircles(lng=location$lon, lat=location$lat, 
                radius=1520, color='red', 
                label=location$address) %>%
     addMarkers(lng=KASSLER$longitude, lat=KASSLER$latitude, 
                label=paste(KASSLER$station, KASSLER$name)) %>%
     addCircles(lng=stations$longitude, lat=stations$latitude, 
                label=paste(stations$station, stations$name)) %>%
     fitBounds(lng1=(FB[1]), lat1=(FB[4]), 
               lng2=(FB[3]), lat2=(FB[2])) 
})
leafletOutput('EC')
```

Column {}
-------------------------------------
###
```{r, echo=FALSE}
stations %>% kbl() %>% kable_styling()
```


<b>Historical</b> {data-icon="ion-stats-bars"}
=====================================  

Column {data-width=200}
-------------------------------------
### <b>station 054452 KASSLER</b>

```{r, echo=FALSE}
  elements %>% kbl() %>% kable_styling()
```

Row {}
-------------------------------------
### <b>Temperature</b>

```{r, echo=FALSE}
  fig <- plot_ly( type = 'scatter', mode = 'lines', showlegend=FALSE ) 
  fig <- fig %>% add_trace(data=historicalWeather, x=~date, y = ~mint, name="min") 
  fig <- fig %>% add_trace(data=historicalWeather, x=~date, y = ~maxt, name="max") 
  fig <- fig %>% layout(yaxis=list(title="Temperature"))
  fig
```

Column {data-width=300}
-------------------------------------
### <b>Precipitation</b>

```{r, echo=FALSE}
  fig <- plot_ly( type = 'scatter', mode = 'lines', showlegend=FALSE ) 
  fig <- fig %>% add_trace(data=historicalWeather, x=~date, y = ~pcpn, name="pcpn") 
  fig <- fig %>% layout(yaxis=list(title="Precipitation"))
  fig
```

### <b>Snowfall</b>

```{r, echo=FALSE}
  fig <- plot_ly( type = 'scatter', mode = 'lines', showlegend=FALSE ) 
  fig <- fig %>% add_trace(data=historicalWeather, x=~date, y = ~snow, name="snow") 
  fig <- fig %>% layout(yaxis=list(title="Snowfall"))
  fig
```

### <b>Snow Depth</b>

```{r, echo=FALSE}
  fig <- plot_ly( type = 'scatter', mode = 'lines', showlegend=FALSE ) 
  fig <- fig %>% add_trace(data=historicalWeather, x=~date, y = ~sndw, name="sndw") 
  fig <- fig %>% layout(yaxis=list(title="Snow Depth"))
  fig
```

### 
<b>
NOAA Online Weather Data (NOWData) 

API ... Applied Climate Information System (ACIS)
https://www.rcc-acis.org/docs_webservices.html
</b>





<b>Forecast</b> {data-icon="ion-stats-bars"}
=====================================  

Column {data-width=150}
-------------------------------------
### <b>Forecast for location</b>
```{r, echo=FALSE}
renderUI({
  HTML(t(RV$init) %>% kbl() %>% kable_styling)
})
```

### <b></b>
```{r, echo=FALSE}
  sliderInput("fc", HTML("<b>when</b>"), min=0, max=14, value=0, step=1)
  actionButton("reset2", "Map")
```

```{r, echo=FALSE}
observeEvent(input$reset2, {
    proxy <- leafletProxy("BC")
    proxy %>% clearShapes() %>%
     addProviderTiles("CartoDB.Positron") %>% 
     setView(lng=RV$init$lon, lat=RV$init$lat, zoom = 9) %>% 
     addCircles(lng=RV$init$lon, lat=RV$init$lat, 
                radius=500, color='red', 
                label=RV$init$address)
})
```

```{r, echo=FALSE}
  observeEvent(input$BC_click, {
  
    RV$init$address <- "map input"
    RV$init$lat <- input$BC_click$lat
    RV$init$lon <- input$BC_click$lng
    
    proxy <- leafletProxy("BC")
    proxy %>% clearShapes() %>% addTiles() %>% 
      setView(lng=RV$init$lon, lat=RV$init$lat, zoom = 9) %>% 
      addCircles(lng=RV$init$lon, lat=RV$init$lat, weight=2, color='red',
                 radius = 500, opacity = 2,
                 highlightOptions = highlightOptions(color = "green", weight = 15),
                 popup=RV$init$address, label=RV$init$address) 
})
```

Column{data-width=350}
-------------------------------------
### <b>Daily forecast from https://api.weather.gov/points/</b>
```{r, echo=FALSE}
renderUI({
    point <- c(RV$init$lat, RV$init$lon)
    RV$url <- GET_forecast_url(point)
    RV$forecastWeather <- GET_forecast_weather(url=RV$url[1,2])
    if ( input$fc > 0 ) {
      HTML(RV$forecastWeather[,c(1,input$fc+1)] %>% 
      kbl(row.names = FALSE) %>% 
      kable_styling)
    }
})
```

Column{}
-------------------------------------
###
```{r, echo=FALSE}
output$BC <- renderLeaflet({

   leaflet() %>% 
     addProviderTiles("CartoDB.Positron") %>% 
     setView(lng=RV$init$lon, lat=RV$init$lat, zoom = 12) %>% 
     addCircles(lng=RV$init$lon, lat=RV$init$lat, 
                radius=520, color='red', 
                label=RV$init$address)
})
leafletOutput('BC')
```

### <b>Hourly forecast from https://api.weather.gov/points/</b>
```{r, echo=FALSE}
renderPlotly({
    
  hourlyForecastWeather <- GET_hourly_forecast_weather(url=RV$url[2,2])
  df <- data.frame(t(hourlyForecastWeather[1,]), 
                   t(hourlyForecastWeather[2,]))
  colnames(df) <- c("time", "temperature")

  if ( input$fc == 0 ) {
    plot_ly(type='scatter', mode='lines', showlegend=FALSE) %>%
      add_trace(data=df, x=~time, y=~temperature, name="forecast",
                line = list(shape = 'spline', smoothing = 5.3))
  } else {
    date1 <- RV$forecastWeather[2,input$fc+1]
    date2 <- RV$forecastWeather[2,input$fc+2]
    w <- df[which(df$time >= date1 & df$time <=date2),]

    plot_ly(type='scatter', mode='lines', showlegend=FALSE) %>%
#      add_trace(data=df, x=~time, y=~temperature, name="forecast",
#                line = list(shape = 'spline', smoothing = 5.3)) %>%
      add_trace(data=w, x=~time, y=~temperature, name="window",
                line = list(shape = 'spline', smoothing = 5.3)) %>%
      add_text(data=w, x=~time, y=~temperature, 
               text=~temperature, textposition = "top")

  }
})
```


<b>forecastGridData</b> {data-icon="ion-stats-bars"}
=====================================  
forecastGridData            https://api.weather.gov/gridpoints/BOU/60,50

Data for a single grid around the selected point


<b>observationStations</b> {data-icon="ion-stats-bars"}
=====================================  
observationStations        https://api.weather.gov/gridpoints/BOU/60,50/stations

A list of weather stations around the selected point.

PLOT 